---
title: "Writing Reports with  R Markdown"
author: "Ernesto Martinez"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, include=FALSE}
library(tidyverse)
library(here)
library(knitr)
```

This report was prepared for the UN. It analyzes the relationship between a country's GPD, life expectancy, and CO2 emissions. Our goal was to determine to what degree a country's economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions will increase with a country's GDP. 

```{r data, include=FALSE, echo=FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))
```



```{r gdp_lifeexp_1997, echo=FALSE}
ggplot(gapminder_1997, aes(x=gdpPercap,y=lifeExp,color=continent,size=pop/1000000)) +
  geom_point() +
  labs(x="GDP per capita",
       y= "Life Expectancy",
       title = "Do people in wealthy countries live longer?",
       size = "Population (in millions)",
       color = "Continent")
```

```{r summary_data, echo=FALSE}
n_countries <- gapminder_1997 %>% 
  select(country) %>%
  n_distinct()

min_gdp <- gapminder_1997 %>%
  select(country,gdpPercap) %>%
  filter(gdpPercap == min(gdpPercap))

max_gdp <- gapminder_1997 %>%
  select(country,gdpPercap) %>%
  filter(gdpPercap == max(gdpPercap))

min_gdp1 <- gapminder_1997 %>%
  select(gdpPercap) %>%
  filter(gdpPercap == min(gdpPercap))

max_gdp1 <- gapminder_1997 %>%
  select(gdpPercap) %>%
  filter(gdpPercap == max(gdpPercap))

gapminder_1997 %>%
  summarise(min=round(min(gdpPercap)))

gapminder_1997 %>%
  summarise(max=round(max(gdpPercap)))
```

The above plot shows the relationship between our GDP per capita and life expectancy for a total of `r n_countries` countries. For this set of countries, economic wealth ranged from a minimum of $`r min_gdp$gdpPercap[1]` in `r min_gdp$country[1]` to a maximum of $`r max_gdp$gdpPercap[1]` per capita in `r max_gdp$country[1]`. 

The above plot shows the relationship between our GDP per capita and life expectancy for a total of `r n_countries` countries. For this set of countries, economic wealth ranged from a minimum of $`r min_gdp1` to a maximum of $`r max_gdp1`. 


| Summary of Data   |
|----------|--------|
| Number of Countries | `r n_countries` |
| Minimum GDP per cap | $`r min_gdp1` |
| Maximum GDP per cap | $`r max_gdp1` |


```{r tables}
gapminder_1997 %>%
  filter(continent == "Oceania") %>%
  kable()
  


```



